\propertitle

### 2019-11-27 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-27T11:44:52+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The crucial point is our belief distribution for the activity of the
sample conditional on the activity of the larger population, given by
formula~\eqref{eq:conditional_hypergeometric} in
appendix~\ref{sec:derivation_maxent_fullnet}, reproduced here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
\end{equation}
which leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our approach rests.
The conditional probability above is characteristic of \enquote{drawing
  without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, consider the pool of all neurons which we believe -- with
equal degree for each neuron -- could have been recorded. $N$ is the size
of that pool. $N$ therefore depends on factors such as: the shape,
dimensions, and technical specifications of the recording probe; the
inaccuracy in the insertion of the probe, leading for example to slightly
different insertion points or angles; the density of neurons around the
probe. But it also depends on the homogeneity of the brain region where the
recording was made: if we believe that it doesn't matter whether the probe
had been inserted in some other point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. Thus $N$ can
only be assessed case by case.

Whether such beliefs are justified, or for which sampling procedures they
can be justified, is a fundamental, deep question of sampling theory, which
we cannot discuss here \citep[see \eg\ the discussions, reviews, and
references in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}.
It is of course possible to probabilistically assess, case by case, whether
this assumption is appropriate. It good to keep in mind that such
assessment rests, in turn, on similar assumptions at a higher level.

@@

If our uncertainty about $N$ spans several orders of magnitude, we can of
course assign a probability to the possible values of $N$ based on our
background information and the frequency data $\yf$. The procedure is
similar to that of \sect~\ref{sec:marginalization} and appendix***. The
probability for the data $\yf$ conditional on a set of constraints $M$ and
$N$ is the exponential of minus the relative entropy~\eqref{eq:relentropy_T}:
\begin{equation}
  \label{eq:prob_data_from_N}
  \pf(\yf \| M, N) = \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
  \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}
If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
\begin{equation}
  \label{eq:prob_N_from_data}
  \pf(N \| M, \yf) \propto
  p(N)\;\exp\bigl\{-T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}

Here is a simple illustrative example with our first data set. Suppose our
uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
to $N=20\,000$. Divide this range roughly into thirds of order of
magnitude, considering the values
$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
of first five moments is sufficient, formula~\eqref{eq:prob_data_from_N}
gives
\begin{equation}
  \label{eq:data1_f_from_N}
  \begin{gathered}
  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
  \pf(\yf \| N=2\,000, M) =  0.00704,\\
  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
  \pf(\yf \| N=10\,000, M) =  0.0150,\\
  \pf(\yf \| N=20\,000, M) =  0.0135.
\end{gathered}
\end{equation}
Since our uncertainty regards a scale factor, it can be represented by
equal pre-data probabilities $1/5$ about these partial orders of magnitude.
Thus from~\eqref{eq:data1_f_from_N} we find
\begin{equation}
  \label{eq:data1_N_from_f}
  \begin{gathered}
  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
  \pf(N=2\,000 \| \yf, M) =  0.140,\\
  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
  \pf(N=10\,000 \| \yf, M) =  0.298,\\
  \pf(N=20\,000 \| \yf, M) =  0.267,
\end{gathered}
\end{equation}
which gives a slightly higher probability to $N=10\,000$.

@@

First of all the sampling relation. We have a set of $N$ units, $\yA$ of
which have some specific property, and we sample in an unknown way $n$ of
the $N$ units. The probability that $\ya$ of the $n$ sampled units have that
property is then given by the hypergeometric distribution
\begin{equation}
  \label{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
\end{equation}
typical of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In the following leave $n$, $N$ implicit in the conditional. If we are
uncertain about the number $\yA$, with belief $P(\yA)$, then by the theorem
of total probability our belief about $\ya$ is
\begin{equation}
  \label{eq:belief_sample_unknown_A}
  p(\ya)= \sum_{\yA} \yG_{\ya\yA} \;P(\yA).
\end{equation}
In our case the units are neurons, the set is the larger population, and the
property is their being active in a specific time bin.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-27T14:35:27+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The crucial point is our belief distribution for the activity of the
sample conditional on the activity of the larger population, given by
formula~\eqref{eq:conditional_hypergeometric} in
appendix~\ref{sec:derivation_maxent_fullnet}, reproduced here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
\end{equation}
leading to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our approach rests.
This conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
ensues when our degree of belief that the $i$th unit (ball or neuron) will
be drawn has the same value for all $i \in \set{1,\dotsc,N}$. Thus,
consider the pool of all neurons which we believe -- with equal degree for
each neuron -- could have been recorded. $N$ is the size of that pool. $N$
therefore depends on factors such as: the shape, dimensions, and technical
specifications of the recording probe; the inaccuracy in the insertion of
the probe, leading for example to slightly different insertion points or
angles; the density of neurons around the probe. But it also depends on the
homogeneity of the brain region where the recording was made: if we believe
that it doesn't matter whether the probe had been inserted in some other
point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
therefore only be assessed case by case.

Whether such equal beliefs are justified, or for which sampling procedures
they can be justified, is the fundamental, deep question of sampling
theory, which we cannot discuss here \citep[see \eg\ the discussions,
reviews, and references
in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}. It is of
course possible to probabilistically assess, case by case, whether this
equality assumption is appropriate. But we must also be aware that such
assessment in turn rests on analogous assumptions at a higher level, and so
on. The probability calculus, just like the logical calculus, cannot yield
conclusions without premisses \citep[p.~182]{johnson1924}.

@@

If our uncertainty about $N$ spans several orders of magnitude we can
assign a probability to the possible values of $N$ based on our background
information and the frequency data $\yf$, similarly to the procedure in
\sect~\ref{sec:marginalization} and appendix~\ref{sec:measure_suff}. The
probability for the data $\yf$ conditional on a set of constraints $M$ and
size $N$ is the exponential of the negative relative
entropy~\eqref{eq:relentropy_T}:
\begin{equation}
  \label{eq:prob_data_from_N}
  \pf(\yf \| M, N) = \binom{T}{T\yf}\;\prod_{a}\pxx(\ya \| M, N)^{T\,\yff_{a}}
  \approx \exp\bigl\{  -T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}
If our pre-data probability for $N$ is $p(N)$, then by Bayes's theorem
\begin{equation}
  \label{eq:prob_N_from_data}
  \pf(N \| M, \yf) \propto
  p(N)\;\exp\bigl\{-T\,\sh[\yf; \pxx(\ya \| M, N)]\bigr\}.
\end{equation}

Here is an illustrative example with our first data set. Suppose our
uncertainty spans slightly more than an order of magnitude, from $N=1\,000$
to $N=20\,000$. Divide this range roughly into thirds of order of
magnitude, considering the values
$N\in \set{1\,000, 2\,000, 5\,000, 10\,000, 20\,000}$. Assuming the set $M$
of first five moments to be sufficient, formula~\eqref{eq:prob_data_from_N}
gives
\begin{equation}
  \label{eq:data1_f_from_N}
  \begin{gathered}
  \pf(\yf \| N=1\,000, M) =  0.00222,\qquad
  \pf(\yf \| N=2\,000, M) =  0.00704,\\
  \pf(\yf \| N=5\,000, M) =  0.0127,\qquad
  \pf(\yf \| N=10\,000, M) =  0.0150,\\
  \pf(\yf \| N=20\,000, M) =  0.0135.
\end{gathered}
\end{equation}
Since our uncertainty regards a scale factor, it can be represented by
equal pre-data probabilities of $1/5$ about these partial orders of
magnitude. Thus from~\eqref{eq:data1_f_from_N} we find
\begin{equation}
  \label{eq:data1_N_from_f}
  \begin{gathered}
  \pf(N=1\,000 \| \yf, M) =  0.044,\qquad
  \pf(N=2\,000 \| \yf, M) =  0.140,\\
  \pf(N=5\,000 \| \yf, M) =  0.251,\qquad
  \pf(N=10\,000 \| \yf, M) =  0.298,\\
  \pf(N=20\,000 \| \yf, M) =  0.267,
\end{gathered}
\end{equation}
which gives a slightly higher probability to $N=10\,000$.

@@

First of all the sampling relation. We have a set of $N$ units, $\yA$ of
which have some specific property, and we sample in an unknown way $n$ of
the $N$ units. The probability that $\ya$ of the $n$ sampled units have that
property is then given by the hypergeometric distribution
\begin{equation}
  \label{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
\end{equation}
typical of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}[\sects~2.1,
3.2]{jeffreys1939_r1983}.
In the following leave $n$, $N$ implicit in the conditional. If we are
uncertain about the number $\yA$, with belief $P(\yA)$, then by the theorem
of total probability our belief about $\ya$ is
\begin{equation}
  \label{eq:belief_sample_unknown_A}
  p(\ya)= \sum_{\yA} \yG_{\ya\yA} \;P(\yA).
\end{equation}
In our case the units are neurons, the set is the larger population, and the
property is their being active in a specific time bin.
